\documentclass[t,handout]{beamer}

\usepackage{beamerthemeshadow}
\usepackage{amsmath} 

\title{The Business Case for a Knowledge Engineer}
\author{Craig Kelly}
\date{\today} 

\begin{document}

\frame{\titlepage} 

\frame{\frametitle{Table of Contents}
    \tableofcontents
} 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The Case}

    \subsection{Data, Knowledge, and Information Management}

    \frame{\frametitle{From Data to Knowledge}
        \begin{itemize}
            \item Drive efficiency by optmizing various business variables like
            pricing, product mix, etc.

            \item Allow limited, but cost-free business expermentation.  Analytical
            models are difficult to builld and verify, but allow reasonable answers
            to what-if's questions from business leaders without changing current
            policies and procedures.

            \item Provide the knowledge necessary for strategic sales.  Analytical
            tools can be developed to help salespeople target new customers and drive
            growth with existing customers.  Non-obvious stratgies might be uncovered,
            such as targeting specific equipment types in certain markets.
        \end{itemize}
    }

    \frame{\frametitle{Ad hoc and IT Handoff}
        \begin{itemize}
            \item The specificity of the job results in sharpened skills and domain
            knowledge necessary to understand and execute complicated ad hoc knowledge
            requests.

            \item Ownership of the question and the answer provide a single touch
            point for the resulting ``semi-automated'' process

            \item Any processes handed off to I.T. have been kept in the ``language
            of IT'' from the beginning of the project, resulting in greatly decreased
            hand-over and implementation times
        \end{itemize}
    }

    \subsection{Business/Market Intelligence}

    \frame{\frametitle{Analysis, Modeling, and Simulation}
        \begin{itemize}
        \item
        In the last thirty years, computer science, statistics, and related fields
        have made huge strides in various techniques for generating intellgence
        that can be leveraged for: strategic decision making, tactical monitoring,
        and new product/market identification and evaluation.

        \item
        Analysis, Modeling, and Simulation of various market conditions
        and processes including the quotation process, fleet dismantling,
        new product introduction, and macro market conditions.
        \end{itemize}
    }

    \frame{\frametitle{Forecasting, Regression, and Clustering}
        \begin{itemize}
        \item
        Forecasting, regression, and machine learning to predict the
        value of various business variables including inquries, sales, repairs,
        and credit default.
        
        \item
        Clustering, comparisons, and link analysis to discover relationships
        between part types, aircraft types, vendor types, seasonality, market
        segments, and other business dimensions
        \end{itemize}
    }

    \subsection{Leverage}

    \frame{\frametitle{Leverage External GE Expertise}
        \begin{itemize}
        \item Looking across GE for best practices, data sources, and tools
        can provide some easy wins

        \item Further examination of currently used external assets, including
        GECAS models and GRC deliverables, could provide extra value not currently
        being utilized
        \end{itemize}
    }

    \frame{\frametitle{Crowdsourcing - Both Internal and External}
        \begin{itemize}
        \item
        Seeking input across all AMS functions for large analytical projects
        would insure any output and deliverables are not unncessarily lopsided.
        In addition, new ideas for analytical projects could be uncovered.

        \item Use external services to provide insight into anonymized data.
        \end{itemize}
    }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Position Description and Requirements}

    \subsection{Business/Domain}

    \frame{\frametitle{Business/Domain}
        Like any knowledge-based job (including I.T. and Finance), a great deal
        of internal business and domain knowledge would be required to effectively
        perform job duties.  However, a more important consideration would be the
        ability of a candidate to leverage the domain expertise extant in the
        Inventory Management department and across the company.
    }

    \subsection{Computational and Mathematical Modeling}

    \frame{\frametitle{Computational and Mathematical Modeling}
        A large part of the position will involve creating, evaluating, applying,
        and updating mathematical models that would form the basis of large
        analytics and projects.  In addition, application of computational modeling
        skills would be needed to complete most analytical projects involving
        ``modern'' simulation and machine learning techniques.
    }

    \subsection{Software Engineering/Data Compliance}

    \frame{\frametitle{Software Engineering/Data Compliance}
        All projects deliverables from this position must maintain compliance
        and be readily handed over to I.T. if they are deemed business critical.
        As a result, standard software engineering and project management skills
        should be utilized to mainatain quality and reduce drag on hand-offs.
    }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Possible Projects}

    \subsection{A Few Sample Projects}

    \frame{\frametitle{A Few Sample Projects}
        Please see the business case for more detail
        
        \begin{description}

        \item[Ad Hoc Analytical Queries]
            Develop system to allow analytical exploration and deep dives without
            impacting ICS transational processing or ``stealing'' I.T. time.

        \item[Portfolio Optimization]
            Using parameterized inputs (from expert what-if's, forecasting, inventory,
            skyline, etc) determine product mix and it's effect on forecasted sales
            versus an optimized product mix.  What are delta's and what are the
            smallest changes to current planning with largest effect?
            
        \item[``Lever'' Modeling]
            Combination of models and simulation to allow exploration of ``levers
            we can pull'', including price elasticity, inventory levels, marketing
            (including controlled inventory listing on ILS)
            
        \end{description}
    }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
       
\end{document}
